Testing Quasar Unification with Clumpy Wind Models

### 2015-12-08 paper_mnras.tex

Here, we continue our tests of  this paradigm by comparing ``average'' quasar spectra to 
synthetic spectra of simple biconical disc wind models, using our state-of-the-art 
Monte Carlo radiative transfer and photoionization code.
In particular, we investigate whether clumping of the outflow allows us to 
produce synthetic spectra in the rest-frame UV that have the characteristics of quasars. 
We find that a simple treatment of clumping (`microclumping') 
allows for a more realistic X-ray luminosities, while maintaining the ionization 
state necessary for strong BAL features in the rest-frame UV. 
We examine the X-ray properties of these simple clumped models

@@

Evidence for dense substructures in AGN winds is widespread.
BALQSOs show complex absorption line profiles \citep{ganguly2006, simonhamann2010}
and exhibit variability in these profile shapes \citep{capellupo2011,capellupo2012,capellupo2014}.
AGN generally show variability in X-ray absorption components \citep[e.g.][]{risaliti2002}
and many models for the BLR consist of clumps embedded in an outflow 
\citep{krolik1981, emmering1992, dekool1995, cassidyraine1996}.
Clumping can be caused by magnetic confinement \citep{dekool1995},
or the instabilities inherent to line-driven winds 
\citep{lucysolomon1970,macgregor1979,carlberg1980,owockirybicki1984,owockirybicki1985}.
Additionally, clumping is required to explain the electron scattering wings of emission lines formed
in line-driven hot star winds \citep{hillier1991eswingsmodel}. Complex substructures 
are also produced in simulations of line-driven 
outflows in AGN, although on very different scales to line-driven instabilities 
\citep{PSK2000,PK04,progakurosawa2010,proga2014}.
Nevertheless, clumpy winds offer an observationally motivated and theoretically 
predicted way to lower the ionization state of a plasma, possibly in tandem
with a shielding scenario.

% An inhomogenous outflow has been proposed as a model for the BLR (REFs)
% Clumping is expected in outflows and BLRs.

% The UV spectrum of {\em non-BAL} QSOs is typified by a series
% of strong emission lines (e.g. \la, \civ, \nv) with an underlying blue continuum
% - the so-called {\sl `big blue bump'} (BBB). The BBB is normally attributed to blackbody-like
% emission from an accretion disc surrounding the central black hole (REF), similar
% to that described by \cite{shakurasunyaev1973}. However,
% a number of issues have arisen relating to this model. First, 
% AGN/QSO spectra exhibit a `break' in the spectrum at around $1000$~\AA 
% which scales only weakly with black hole mass. This potentially 
% suggests a problem with a thin disc model (e.g. Antonucci). 
% However, it is possible that this is the result of incorrect intergalactic medium
% (IGM) corrections (REF) or corresponds to the temperature 
% at which a line-driven wind carries mass away from the system (laor davis).
% Second, results from microlensing (REFs) imply that the disc emission 
% region is $\sim4$ times larger than one might expect from a Shakura-Sunyaev
% model. Inhomogenous discs have been proposed as an alternative (REFs).
% These observations appear to pose problems for the thin disc model of luminous,
% sub-Eddington AGN. However, recent results from \cite{capellupo2015} find 
% that if one includes a combination of mass-loss, general relativity (GR) and Comptonisation
% then AGN SEDs can, in general, be fitted well with accretion disc models.
% Uncovering the intrinsic disc SED and understanding the effect of the outflow on the accretion 
% mechanism is clearly crucial if we are to properly understand the physical nature
% of AGN.

% The geometry and size of the BLR is also a matter of contention. 
% The main constraints on the emission region size come from microlensing (REFs)
% and reverberation mapping (REFs). While these observations have mostly been
% carried out for Seyfert galaxies, there are also a few instances of these methods
% being applied to quasars (REFs). A number of different proposals have been made 
% for the origin of the BLR. Early AGN unification scenarios posited that the broad emission lines
% where produced by clouds of plasma orbiting fairly close to the BH (refs).
% Since then, multiple interpretations of a disc wind model have been proposed,
% with varying radii and geometries (refs).

We have been engaged in a project to determine whether it is possible to simulate the properties of the spectra of AGN, including BAL QSOs, using simple kinematic prescriptions for biconical disc winds using a Monte Carlo radiative transfer (MCRT) code that calculates the ionization structure of the wind and simulates the spectra from such a system \citep[][hereafter H13]{simlong2008,sim2010,higginbottom2013}.  The results have encouraging in the sense that in H13, we showed we could produce simulated spectra that resembled that of BAL QSOs, as long as the luminosity of the central engine was relatively low, of order \POW{43}{\LUM} and the mass loss rate was relatively high, of order the mass accretion rate.  However, at higher luminosities, the wind was so ionized that UV absorption lines were not produced.  In addition, and in part due to limitations in our radiative transfer code, the model failed to produce spectra with strong emission lines at any inclination angle.

@@

For this study, we use the MCRT code \py\ we have developed to carry out our 
radiative transfer and photoionization simulations in non-local thermodynamic
equilibrium (non-LTE). The code can be model a variety of
disc-wind systems; it has been used with application to accreting white dwarfs 
(Long \& Knigge 2002, hereafter LK02; Noebauer et al. 2010; 
Matthews et al. 2015, hereafter M15), young-stellar objects 
\citep{simmacro2005} and quasars/AGN (H13, H14).\nocite{noebauer, M15, LK02}

The basic premise involves generating photons from an accretion disc
and central object, and computing their radiative transfer through an outflow.
This outflow is discretized into $n_x \times n_z$ cells in a 2.5D
cylindrical geometry with azimuthal symmetry.
First, the code gradually converges on the global temperature and ionization structure.
Once these ``ionization cycles'' are complete, then a series of further iterations
allow one to obtain a detailed synthetic spectrum.
LK02 provides the basic description of the original code; improvements, 
see especially \cite{simmacro2005}, H13 and M15, have been 
described in subsequent reports.  
%A full description of the code will be published elsewhere; 
We focus here on the specific changes made to improve the ionization calculation 
and to allow for clumping in the wind.
%These studies contain extensive detail on the code,  so we only briefly describe the key elements of the global ionization calculation and other important aspects.

@@

Our approach to line transfer is based upon the macro-atom implementation developed by 
\cite{lucy2002, lucy2003}, in which the energy flows through the system are described in 
terms of indivisible energy quanta of radiant or kinetic energy 
(`$r-$packets' and `$k-$packets' respectively; see also section~\ref{sec:photon_sources}). 
In our case, for reasons of computational efficiency, we adopt the  hybrid macro-atom scheme 
described by M15.
%(an improvement over that used by H14 for our earlier quasar study), 
In this scheme, the energy packets interact with either two-level `simple ions' 
or full `macro-atoms'. This allows one to treat non-LTE line transfer in radiative equilibrium
without approximation for elements that are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple-ions (see M15). In this study,
only H is treated as a macro-atom, because 
we expect recombination to be important
in determining their level populations and resultant line emission, and because we were especially interested in the contribution to AGN spectra of \LA.  
H13 treated all atoms in a two-level approximation.

\subsection{Ionization Scheme}

@@

In order to calculate the photoionization rate, 
we model the SED in a grid cell using the technique described by H13. In this scheme,
the mean intensity, $J_{\nu}$ in a series of $n$ bands is modeled as either a power law or exponential
in frequency $\nu$, with the fit parameters deduced from band-limited radiation field estimators.
The ionization rate out of ion $j$ can then be written as 
\begin{equation}
R_{j,j+1}(J)= 
\displaystyle{
n_j \left(C_{j} n_e + 
\sum_{band~i=0}^{n}~
{\int_{\nu_i}^{~\nu_i+1} \frac{4 \pi J_{\nu,i}\sigma_j(\nu)}  {h \nu} d\nu}
\right),}
\end{equation}
where $\sigma_j$ is the photoionization cross-section, $n_e$
is the electron density and $C_{j}$
represents the collisional ionization coefficient.
The recombination rate {\em into} ion $j$ is given by 
\begin{equation}
R_{j+1,j}(T_e) = (\alpha^j_{RR} + \alpha^j_{CR}) n_{j+1} n_e,
\end{equation}
where each $\alpha^j$ here is the recombination rate coefficient into the ground state of ion $j$.
The subscripts denote radiative and 
collisional (three-body) recombination, respectively. 
$T_e$ is the electron temperature.
%% We neglect recombination to and from excited states in the simple-ion calculation. 
For simple-ions, we use a dilute Boltzmann equation to calculate 
the population of level $k$ in ionic stage $j$,
\begin{equation}
\frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
\end{equation}
Here $z_j$ is the partition function of ionic stage $j$,
$T_R$ is the effective radiation temperature,
$E_k$ is the energy difference between level $k$ and the ground state,
and $g_k$ is the statistical weight of level $k$. 
We stress that this approximation is not required for ions
treated as macro-atoms.

@@

We include all free-free, bound-free and bound-bound heating
and cooling processes in the model. For radiative transfer purposes
we treat electron scattering in the Thomson limit, 
but take full account of Compton heating and cooling when
calculating the thermal balance of the plasma (see H13).
Adiabatic cooling is included and represents the only
departure from strict radiative equilibrium, but is insignificant 
in most of the outflow.

@@

We include two sources of r-packets in our model:
An accretion disc and central X-ray source.
The accretion disc is assumed to be geometrically thin, but optically thick.
We thus treat the disc as an ensemble of blackbodies with a 
\cite{shakurasunyaev1973} effective temperature profile. 
The emergent SED is then determined by the specified accretion rate ($\dot{m}$)
and central BH mass ($M_{BH}$).
All photon sources in our model are opaque, meaning
that r-packets which strike them are destroyed.
The inner radius of the disc extends to the innermost 
stable circular orbit (ISCO) of the BH. 
We assume a Schwarzchild BH with an ISCO at $6~r_G$, where 
$r_G = GM_{BH}/c^2$ is the gravitational radius.
For a $10^9~M_\odot$ black hole, this is equal to $8.8\times10^{14}~{\rm cm}$ 
or $\sim10^{-4}~{\rm pc}$.

The X-ray source is treated as an isotropic sphere at the ISCO,
which emits r-packets according to a power law in flux with index $\alpha_X$, of the form
\begin{equation}
F_X (\nu) = K_X \nu^{\alpha_X}.
\end{equation}
The normalisation, $K_X$ of this power law is such that it 
produces the specified 2-10~keV luminosity, $L_X$.
In addition to the disc and X-ray source, 
the wind is able to reprocess radiation. However, new 
photon packets are not produced in the wind (as in LK02). 
Instead, this reprocessing is dealt with by enforcing strict
radiative equilibrium ({\em modulo} adiabatic cooling; see section~2.3)
via an indivisible energy packet
constraint (see Lucy 2002, M15).

@@

\subsection{A First Approximation for Clumping}

Our previous modelling efforts assumed a smooth outflow, 
in which the density at a given point was determined only by the 
kinematic parameters and mass loss rate. However, as already discussed,
AGN winds exhibit significant substructure -- the outflow is expected to be
{\em clumpy}, rather than smooth, and probably on a variety of scales. 
Implementing a treatment of clumping is challenging, for
two main reasons. First, there are significant 
computational difficulties associated with adequately
resolving and realistically modelling a series of small scale, high density
regions with a MCRT code. Second, the addition of multiple additional degrees of
freedom in the model results in significantly wider parameter space.
Unfortunately, the physical scale lengths and density contrasts 
associated with these parameters cannot be well-constrained from observations.

To allow for clumping in our outflow we adopt a simple approximation
used successfully in stellar wind modelling, known as 
{\em microclumping} \citep{hamann1998,hamann2008}(MORE REFS). 
The key assumption is that typical clump sizes are much smaller than the 
typical photon mean free path, and thus the clumps are 
both geometrically and optically thin. This approach 
allows one to introduce a `volume filling factor', $f_V$. 
The intra-clump medium is assumed to be a vacuum, so the 
density of the clumps is then multiplied by the ``density enhancement'' 
$D=1/f_V$. Opacities, $\kappa$, and emissivities, $\epsilon$, 
can then be expressed as 
\begin{equation}
\kappa = f_V \kappa_C(D);~~\epsilon = f_V \epsilon_C(D).
\end{equation}
Here the subscript $C$ denotes that the quantity is calculated using the 
enhanced density in the clump. The resultant effect is that {\em linear} processes,
such as electron scattering, are unchanged, as $f_V$ and $D$ will cancel out. 
However, any quantity which scales with the {\em square} of density, 
such as collisional excitation or recombination, will increase by a factor of $D$.

Clumping the wind has an important effect on the ionization state and has
been proposed as a solution to the so-called `over-ionization problem' in 
disc winds \citep{hamann2013}. This is the main motivation for incorporating microclumping
into our model. This treatment is first-order; it does not adequately
represent the complex substructures and stratifications in ionization
state we expect in AGN outflows. Nevertheless, clumping is clearly
important in these flows, and this parameterization allows a simple estimate
for the effect clumping might have on the ionization state and emergent 
line emission. 
% It is also encouraging that microclumping has been used 
% successfully in fits to O-star wind spectra \citep{hillier1991eswingsmodel}.

@@

Using this prescription, we conducted a limited parameter
search over a 5-dimensional parameter space involving the 
variables $r_{min}$, $\theta_{min}$, $f_V$, $\alpha$ and $R_v$.
The grid points are shown in table 1.
The aim here was to first fix $M_{BH}$ and $\dot{m}$ to their H13 values,
and increase $L_X$ to $10^{45}$~erg~s$^{-1}$ (a more realistic value for a 
quasar of $10^9M_\odot$ and an eddington fraction of $0.2$; see section~\ref{sec:xray}). 
We then evaluated these models based on the following criteria:
\begin{itemize}
\item Does the model avoid over-ionization and thus produce UV absorption lines 
with $BI > 0$ at $\sim20\%$ of viewing angles?
\item Does significant line emission emerge at low inclinations, with $EW\sim40\AA$ in \civ?
\item Do H recombination lines appear in the spectrum, $EW\sim50\AA$ in \la?
\item Do a small subset of BAL angles produce LoBAL features?
\item Does the model compare favourably to quasar composite spectra?
\end{itemize}
In this section, we present one of the most promising models and discuss
the various successes and failures with respect to the above criteria.
This allows us to gain insight into fundamental geometrical 
and physical constraints and assess the potential for unification. 
The full grid, including output synthetic spectra and plots can be found at
\url{jhmatthews.github.io/quasar-wind-grid/}.

@@

\begin{table}
\begin{tabular}{p{3cm}p{4cm}}
\hline Next-generation Model Parameters 	&	 Value \\ 
\hline \hline 
$M_{BH}$ 	 &	 $1\times 10^9~\rm{M_{\odot}}$ \\ 
$\dot{M}_{acc}$ 	 &	 $5~M_{\odot}yr^{-1} \simeq 0.2~\dot{M}_{Edd}$\\ 
$\alpha_X$ 	 &	 $-0.9$ \\ 
$L_{X} $ 	 &	 $10^{45}~\rm{ergs~s^{-1}}$$^*$ \\ 
$r_{disc}(min)=r_{X}$   &	 $6r_g=8.8\times10^{14}~{\rm cm}$ \\ 
$r_{disc}(max)$   &	 $3400r_g = 5\times10^{17}~{\rm cm}$ \\ 
$\dot{M}_{wind}$  &	 $5~M_{\odot}yr^{-1}$ \\ 
$r_{min}$ 	&	 $300r_{g} = 4.4\times10^{16}~{\rm cm}$\\ 
$r_{max}$ 	&	 $600r_{g} = 8.8\times10^{16}~{\rm cm}$ \\ 
$\theta_{min}$ 	&	 $70.0^{\circ}$ \\ 
$\theta_{max}$ 	&	 $82.0^{\circ}$ \\ 
$\lambda$ 	&	 $0$ \\ 
$v_{\infty}(r_0)$ 	&	 $v_{esc}(r_0)$ \\ 
$R_v$  	        &	 $10^{19}$cm$^*$ \\ 
$\alpha$ 	&	 $0.6^*$ \\
$f_V$ 	&	 $0.01^*$  \\
$n_x$ 	&	 $100$  \\
$n_z$ 	&	 $200$  \\
\hline 
% Derived Parameters 	&	 Value \\ 
% \hline \hline
% $L_{\nu}(2500\mbox{\scriptsize{\AA}})$&	 $6.3\times10^{30}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{\nu}(2keV)$   &	 $1.2\times10^{25}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{bol}$ 	 &	 $2.4\times10^{46}~\rm{ergs~s^{-1}}$\\
% $M_{bol}$ 	 &	 -27.4\\ 
% $M_u$            &	 -26.2\\ 
% $\alpha_{OX} $ 	 &	 -2.2\\ 
\end{tabular}
\caption{Wind geometry parameters used in the model.}
\label{wind_param}
\end{table}

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/uvspec_estfix.png}
\caption
{
Synthetic spectra at five viewing angles in our model. The coloured lines
show different quasar and BAL quasar composites, and the dotted line shows a disc
only continuum to show the effect of the outflow on the continuum level.  
HST STIS spectrum of PG0946+301 (Arav et al. 2000)
}
\label{fig:uvspec}
\end{figure*} %fullpage

>>>>

Here, we continue our tests of  this paradigm by comparing quasar spectra to 
synthetic spectra of simple biconical disc wind models, produced with our state-of-the-art 
Monte Carlo radiative transfer and photoionization code PYTHON.  Previously, we have shown that we could produce synthetic spectra resembling those of observed AGN for plausible wind models, but only if the X-ray luminosity was limited to about  10$^{43}$ erg s$^{-1}$.  Here , we investigate the  degree to which  clumping of the outflow allows us to 
produce synthetic spectra in the rest-frame UV that have the characteristics of quasars with X-ray luminosities as large as 10$^{45}$ erg s$^{-1}$. 
We conclude, perhaps not surprisingly,  that clumping does maintain the ionization state of the wind necessary for strong BAL features in the rest-frame UV at more realistic X-ray luminosities. We examine the X-ray properties of these simple clumped models

@@

Evidence for dense substructures in AGN winds is widespread.
BALQSOs show complex absorption line profiles \citep{ganguly2006, simonhamann2010}
and exhibit variability in these profile shapes \citep{capellupo2011,capellupo2012,capellupo2014}.
AGN generally show variability in X-ray absorption components \citep[e.g.][]{risaliti2002}
and many models for the BLR consist of clumps embedded in an outflow 
\citep{krolik1981, emmering1992, dekool1995, cassidyraine1996}.
Clumping can be caused by magnetic confinement \citep{dekool1995},
or the instabilities inherent to line-driven winds 
\citep{lucysolomon1970,macgregor1979,carlberg1980,owockirybicki1984,owockirybicki1985}.
Additionally, clumping is required to explain the electron scattering wings of emission lines formed
in line-driven hot star winds \citep{hillier1991eswingsmodel}. Complex substructures 
are also produced in simulations of line-driven 
outflows in AGN, although on very different scales to line-driven instabilities 
\citep{PSK2000,PK04,progakurosawa2010,proga2014}.  
% {\bf KSL: I do not believe the AGN simulations have the resolution to produce clumping on the scale of hot star winds.
% JM: that's why I mention scales, although I can specifically mention resolution if we
% want.}
Nevertheless, clumpy winds offer an observationally motivated and theoretically 
predicted way to lower the ionization state of a plasma, possibly in tandem
with a shielding scenario.

We have been engaged in a project to determine whether it is possible to simulate the properties of the spectra of AGN, including BALQSOs, using simple kinematic prescriptions for biconical disc winds using a Monte Carlo radiative transfer (MCRT) code that calculates the ionization structure of the wind and simulates the spectra from such a system \citep[][hereafter H13]{simlong2008,sim2010,higginbottom2013}.  The results have been encouraging in the sense that in H13, we showed we could produce simulated spectra that resembled that of BALQSOs, as long as the luminosity of the X-ray source was relatively low, of order \POW{43}{\LUM} and the mass loss rate was relatively high, of order the mass accretion rate.  However, at higher X-ray luminosities, the wind was so ionized that UV absorption lines were not produced.  In addition, and in part due to limitations in our radiative transfer code, the model failed to produce spectra with strong emission lines at any inclination angle.

@@

For this study, we use the MCRT code \py 
\footnote{Named {\sl c. 1995}, predating the inexorable rise of a certain progamming language.} we have developed to carry out our 
radiative transfer and photoionization simulations in non-local-thermodynamic-equilibrium 
(non-LTE). The code can be used to model a variety of
disc-wind systems; it has been used with application to accreting white dwarfs 
(Long \& Knigge 2002, hereafter LK02; Noebauer et al. 2010; 
Matthews et al. 2015, hereafter M15), young-stellar objects 
\citep{simmacro2005} and quasars/AGN (H13, H14).\nocite{noebauer, M15, LK02}

The code operates as follows:   This outflow is discretized into $n_x \times n_z$ cells in a 2.5D
cylindrical geometry with azimuthal symmetry. From some initial conditions in each cell, the code first calculates the ionization structure of the wind in a series of iterations.  Each iteration in an ``ionization cycle'' consists of generating photons, actually photon packets, from an accretion disc
and central object, and calculating how these photon bundles scatter through the wind (eventually escaping the outflow or hitting the disk). Then updating the ionization structure based on the properties of the radiation field in each cell, and the process is repeated.   Once the ionization structure has converged, the ionization structure is held fixed, and  synthetic spectra are generated at specific inclination angles in a series of "spectral cycles". LK02 provide a more detailed  description of the original code; various improvements have been made since then and are described by \cite{simmacro2005}, H13 and M15.  
%A full description of the code will be published elsewhere; 
We focus here on the specific changes made for this study intended to improve the ionization calculation of H and He
and to allow for clumping in the wind.
%These studies contain extensive detail on the code,  so we only briefly describe the key elements of the global ionization calculation and other important aspects.

@@

Our approach to line transfer is based upon the macro-atom implementation developed by 
\cite{lucy2002, lucy2003}, in which the energy flows through the system are described in 
terms of indivisible energy quanta of radiant or kinetic energy 
(`$r-$packets' and `$k-$packets' respectively; see also section~\ref{sec:photon_sources}). 
In our case, for reasons of computational efficiency, we adopt the  hybrid macro-atom scheme 
described by M15.
%(an improvement over that used by H14 for our earlier quasar study), 
In this scheme, the energy packets interact with either two-level `simple ions' 
or full `macro-atoms'. This allows one to treat non-LTE line transfer in radiative equilibrium, which assumes both statistical equilibrium and that 
radiative heating balances radiative cooling.
% {\bf KSL: what does "in radiative equilibrium" here mean? 
% JM: Defined.}
without approximation for elements that are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple-ions (see M15). In this study,
only H and He are treated as a macro-atom, because 
we expect recombination to be important
in determining their level populations and resultant line emission, 
and because we are especially interested in the contribution to 
AGN spectra of \LA.  
H13 treated all atoms in a two-level approximation.  
% {\bf KSL:
% I thought we were also doing He. JM: yep, corrected.}

\subsection{Ionization treatment}

@@

In order to calculate the photoionization rate, 
we model the SED in a grid cell using the technique described by H13. In this scheme,
the mean intensity, $J_{\nu}$ in a series of $n$ bands is modeled as either a power law or exponential
in frequency $\nu$, with the fit parameters deduced from band-limited radiation field estimators.
This allows the calculation of a photoionization rate estimator. Ion abundances are
then calculated by solving the rate equations between ions. We include collisional ionization
and photoionization balanced with radiative, 
dielectronic and collisional (three-body) recombination.
As in M15, we use a dilute Boltzmann approximation to calculate 
the population of levels for simple-ions. We stress that this approximation 
is not required for ions treated as macro-atoms. 
% \begin{equation}
% J_{\nu,i}=K_{pl}\nu^{\alpha_{pl}},
% \end{equation}
% for a band $i$, or an exponential 
% \begin{equation}
% J_{\nu,i}=K_{exp} e^{(-h\nu/k T_{exp})}.
% \end{equation}
% Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
% deduced from the band-limited radiation field estimators.
% The ionization rate out of ion $j$ can then be written as 
% \begin{equation}
% R_{j,j+1}(J)= 
% \displaystyle{
% n_j \left(C_{j} n_e + 
% \sum_{band~i=0}^{n}~
% {\int_{\nu_i}^{~\nu_i+1} \frac{4 \pi J_{\nu,i}\sigma_j(\nu)}  {h \nu} d\nu}
% \right),}
% \end{equation}
% where $\sigma_j$ is the photoionization cross-section, $n_e$
% is the electron density and $C_{j}$
% represents the collisional ionization coefficient.
% The recombination rate {\em into} ion $j$ is given by 
% \begin{equation}
% R_{j+1,j}(T_e) = (\alpha^j_{RR} + \alpha^j_{CR}) n_{j+1} n_e,
% \end{equation}
% where each $\alpha^j$ here is the recombination rate coefficient into the ground state of ion $j$.
% The subscripts denote radiative and 
% collisional (three-body) recombination, respectively. 
% $T_e$ is the electron temperature.
% %% We neglect recombination to and from excited states in the simple-ion calculation. 
% For simple-ions, we use a dilute Boltzmann approximation to calculate 
% the population of level $k$ in ionic stage $j$,
% \begin{equation}
% \frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
% \end{equation}
% Here $z_j$ is the partition function of ionic stage $j$,
% $T_R$ is the effective radiation temperature,
% $E_k$ is the energy difference between level $k$ and the ground state,
% and $g_k$ is the statistical weight of level $k$. 
% We stress that this approximation is not required for ions
% treated as macro-atoms. 
% {\bf KSL: 
% It seems a bit odd that you explain what we do with simple ions our old approach, but don't say what is done for macro-atoms,.  You only say it is not the other.
% JM: We've shortened the section on simple-ions as agreed..}

@@

We include  free-free, bound-free and bound-bound heating
and cooling processes in the model. For radiative transfer purposes
we treat electron scattering in the Thomson limit, 
but take full account of Compton heating and cooling when
calculating the thermal balance of the plasma (see H13).
Adiabatic cooling is included, but is insignificant 
in most of the outflow.

@@

We include two sources of r-packets in our model:
An accretion disc and central X-ray source.
The accretion disc is assumed to be geometrically thin, but optically thick.
Accordingly, we treat the disc as an ensemble of blackbodies with a 
\cite{shakurasunyaev1973} effective temperature profile. 
The emergent SED is then determined by the specified accretion rate ($\dot{m}$)
and central BH mass ($M_{BH}$).
All photon sources in our model are opaque, meaning
that r-packets that strike them are destroyed.
The inner radius of the disc extends to the innermost 
stable circular orbit (ISCO) of the BH. 
We assume a Schwarzchild BH with an ISCO at $6~r_G$, where 
$r_G = GM_{BH}/c^2$ is the gravitational radius.
For a $10^9~M_\odot$ black hole, this is equal to $8.8\times10^{14}~{\rm cm}$ 
or $\sim10^{-4}~{\rm pc}$.

The X-ray source is treated as an isotropic sphere at the ISCO,
which emits r-packets according to a power law in flux with index $\alpha_X$, of the form
\begin{equation}
F_X (\nu) = K_X \nu^{\alpha_X}.
\end{equation}
The normalisation, $K_X$ of this power law is such that it 
produces the specified 2-10~keV luminosity, $L_X$.
In addition to the disc and X-ray source, 
the wind is able to reprocess radiation. However, new 
photon packets are not produced in the wind (as in LK02). 
Instead, this reprocessing is dealt with by enforcing strict
radiative equilibrium ({\em modulo} adiabatic cooling; see section~2.3)
via an indivisible energy packet
constraint (see Lucy 2002, M15).  
% {\bf I confess, I do not understand what is being said here.  I would like to understand it better, and I would like it explained better.  How do we avoid, generating for example recombination photons from non-macro atoms.}

@@

\subsection{A Simple Approximation for Clumping}

Our previous modelling efforts, we have assumed a smooth outflow, 
in which the density at a given point was determined only by the 
kinematic parameters and mass loss rate. However, as already discussed,
AGN winds exhibit significant substructure -- the outflow is expected to be
{\em clumpy}, rather than smooth, and probably on a variety of scales. 
Implementing a treatment of clumping is challenging, for
two main reasons. First, the physical scale lengths and density contrasts 
associated with these parameters are not well-constrained from observations.  
Second, there are significant computational difficulties associated with adequately
resolving and realistically modelling a series of small scale, high density
regions with a MCRT code. 
% Second, the addition of multiple additional degrees of
% freedom in the model results in significantly wider parameter space.
% Unfortunately, the physical scale lengths and density contrasts 
% associated with these parameters are not well-constrained from observations.  
% {\bf KSL: Reverse the order of these difficulties.  The most important is that it is not clear what kind of clumping an AGN wind has; the computational problem is a secondary one JM: Done..}

Given the lack of knowledge about the actual type of clumping, we have adopt a simple approximation
used successfully in stellar wind modelling, known as 
{\em microclumping} \citep{hamann1998,hamann2008}(MORE REFS), 
which minimizes the computational difficulties. 
The underlying assumption in microclumping is that clump sizes are much smaller than the 
typical photon mean free path, and thus the clumps are 
both geometrically and optically thin. This approach 
allows one to introduce a `volume filling factor', $f_V$. 
The intra-clump medium is assumed to be a vacuum, so the 
density of the clumps is then multiplied by the ``density enhancement'' 
$D=1/f_V$. Opacities, $\kappa$, and emissivities, $\epsilon$, 
can then be expressed as 
\begin{equation}
\kappa = f_V \kappa_C(D);~~\epsilon = f_V \epsilon_C(D).
\end{equation}
Here the subscript $C$ denotes that the quantity is calculated using the 
enhanced density in the clump. The resultant effect is that, {\em for fixed temperature},
 processes that are {\em linear} in density, such as electron scattering, are unchanged, 
as $f_V$ and $D$ will cancel out. However, any quantity that scales with the {\em square} of density, 
such as collisional excitation or recombination, will increase by a factor of $D$.
However, in the event that one of these

Clumping the wind has an important effect on the ionization state and has
been proposed as a solution to the so-called `over-ionization problem' in 
disc winds \citep{hamann2013}. This is the main motivation for incorporating microclumping
into our model. This treatment is necessarily simple; it does not adequately
represent the complex substructures and stratifications in ionization
state we expect in AGN outflows. Nevertheless,  this parameterization allows simple estimates
of the effect clumping might have on the ionization state and emergent 
line emission. 
% It is also encouraging that microclumping has been used 
% successfully in fits to O-star wind spectra \citep{hillier1991eswingsmodel}.

@@

Using this prescription, we conducted a limited parameter
search over a 5-dimensional parameter space involving the 
variables $r_{min}$, $\theta_{min}$, $f_V$, $\alpha$ and $R_v$.
The grid points are shown in table 1.
The aim here was to first fix $M_{BH}$ and $\dot{m}$ to their H13 values,
and increase $L_X$ to $10^{45}$~erg~s$^{-1}$ (a more realistic value for a 
quasar of $10^9M_\odot$ and an eddington fraction of $0.2$; see section~\ref{sec:xray}). 
We then evaluated these models based on the following criteria:
\begin{itemize}
\item Does the model avoid over-ionization and thus produce UV absorption lines 
with $BI > 0$ at $\sim20\%$ of viewing angles?
\item Do the synthetic spectra show  line emission emerge at low inclinations, with $EW\sim40\AA$ in \civ?
\item Do H recombination lines appear in the spectra, $EW\sim50\AA$ in \la?
\item Do the spectra  produce LoBAL features at a  small subset of BAL angles?
\item Does the model spectra resemble quasar composite spectra?
\end{itemize}
In the next section, we present one of the most promising models and discuss
the various successes and failures with respect to the above criteria.
This allows us to gain insight into fundamental geometrical 
and physical constraints and assess the potential for unification. 
We then discuss the sensitivity to key parameters in section 5.
The full grid, including output synthetic spectra and plots can be found at
\url{jhmatthews.github.io/quasar-wind-grid/}.

% {\bf I still think we need to avoid using model to be a generic term to describe the outflow and the spectra.  If you mean the spectra use "model spectra" or "synthetic spectra'.  Note - the whole question of the model grid and the selection of one specific instantiation is not dealt with well.}

@@

\begin{table}
\begin{tabular}{p{3cm}p{4cm}}
\hline Next-generation Model Parameters 	&	 Value \\ 
\hline \hline 
$M_{BH}$ 	 &	 $1\times 10^9~\rm{M_{\odot}}$ \\ 
$\dot{M}_{acc}$ 	 &	 $5~M_{\odot}yr^{-1} \simeq 0.2~\dot{M}_{Edd}$\\ 
$\alpha_X$ 	 &	 $-0.9$ \\ 
$L_{X} $ 	 &	 $10^{45}~\rm{erg~s^{-1}}$$^*$ \\ 
$r_{disc}(min)=r_{X}$   &	 $6r_g=8.8\times10^{14}~{\rm cm}$ \\ 
$r_{disc}(max)$   &	 $3400r_g = 5\times10^{17}~{\rm cm}$ \\ 
$\dot{M}_{wind}$  &	 $5~M_{\odot}yr^{-1}$ \\ 
$r_{min}$ 	&	 $300r_{g} = 4.4\times10^{16}~{\rm cm}$\\ 
$r_{max}$ 	&	 $600r_{g} = 8.8\times10^{16}~{\rm cm}$ \\ 
$\theta_{min}$ 	&	 $70.0^{\circ}$ \\ 
$\theta_{max}$ 	&	 $82.0^{\circ}$ \\ 
$\lambda$ 	&	 $0$ \\ 
$v_{\infty}(r_0)$ 	&	 $v_{esc}(r_0)$ \\ 
$R_v$  	        &	 $10^{19}$cm$^*$ \\ 
$\alpha$ 	&	 $0.6^*$ \\
$f_V$ 	&	 $0.01^*$  \\
$n_x$ 	&	 $100$  \\
$n_z$ 	&	 $200$  \\
\hline 
% Derived Parameters 	&	 Value \\ 
% \hline \hline
% $L_{\nu}(2500\mbox{\scriptsize{\AA}})$&	 $6.3\times10^{30}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{\nu}(2keV)$   &	 $1.2\times10^{25}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{bol}$ 	 &	 $2.4\times10^{46}~\rm{ergs~s^{-1}}$\\
% $M_{bol}$ 	 &	 -27.4\\ 
% $M_u$            &	 -26.2\\ 
% $\alpha_{OX} $ 	 &	 -2.2\\ 
\end{tabular}
\caption{Wind geometry parameters 
used in the model, as defined in the text and figure 1.
Parameters differing from the benchmark model of H13 are 
highlighted with an asterisk.}
\label{wind_param}
\end{table}

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/uvspec_new.png}
\caption
{
Synthetic spectra at four viewing angles in our model. At 
$20^\circ$ and $60^\circ$ we show a comparison to an SDSS quasar composite
from Recihard et al. (2003). At $73^\circ$ and $76^\circ$ we show a comparison to
an {\sl HST} STIS spectrum of the high BALnicity BALQSO 
PG0946+301 (Arav et al. 2000), and an SDSS spectrum of the LoBAL quasar 
XXXXX (REF), respectively. The dotted line shows a disc
only continuum to show the effect of the outflow on the continuum level. 
All the spectra are scaled to the model flux at $2000$\AA, expect for the 
{\sl HST} STIS spectrum of PG0946+301, which is scaled using a continuum fit
due to the incomplete wavelength coverage.
% HST STIS spectrum of PG0946+301 (Arav et al. 2000)
}
\label{fig:uvspec}
\end{figure*} %fullpage
